---
title: "hw2"
author: "Brandon Xicon;xiconb"
date: "2/3/2020"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This homework is about syntax and basic plotting functions in R, and is **due by the end of Feb 7**. 

You can use the RMarkDown file with these homework instructions as a template for your work. 
Make sure to replace "Your Name" and "your_github_username" in the YAML with your name and Github username.

**Submission instructions**: Create a folder named `ds202_hw2`, and name the RMarkDown file including your solutions `hw2.Rmd` under this folder. For submission, create a GitHub repository named `ds202_hw2` under your account, and push both `hw2.Rmd` and the knitted `hw2.html` to GitHub before the deadline. I will assume you use the same GitHub username as for your HW1 submission. The latest version of your homework appearing on GitHub before the deadline is graded. *It is crucial to follow the exact folder structure and file names*, so that your homework can be reproduced and graded by an automated script. Some questions will require the following answer format.

## Example question and answer
0\. What is the smallest positive odd integer? Storage it in a variable named `a0` and print it.

**Example answer**:
In the RMarkDown file, type in
<pre><code>```{r}
a0 <- 1
a0
```</code></pre>
The knitted html file will show:
```{r}
a0 <- 1
a0
```

## R commands
1. Today's temperature has a high of 24 and a low of 15 degree fahrenheit.  Find the corresponding degress in celsius and storage it in a (single) vector named `cel`. Print the variable.
```{r}
cel<-c(-4.44,-9.44)
cel
```
2. Create a vector of length 200. The odd entries should be 3 and even entries should be 1. Storage the created in a variable named `vec200` and print.
```{r}
vec200<-c(1,3)
for(i in 3:200)
{
if(i%%2){
vec200<-c(vec200,1)
}
else{
vec200<-c(vec200,3)
}
}
vec200
```
3. How many perfect squares are there between 1 and 2001? Storage the answer in a  variable named `numPerfect` and print. You may want to look at `?round`.
```{r}
numperfect<-trunc(sqrt(2001))
numperfect

```

## Cars and mileage
For this part we will use the `mpg` data set, you can access the data by first installing (using `install.packages('ggplot2')`) and then loading the package `ggplot2` (using `library(ggplot2)`) into your R session. The `mpg` data includes information about fuel economy of popular car models
for the years 1999 and 2008 as collected by the US Environmental Protection Agency. An answer should include a short paragraph of text for the explanation to a (human) reader, together with appropriate R code, outputs, and graphs.

Most exercises are from the ggplot2 book section 2.2.1. You might find it helpful for working on the homework to read through this section. *However, all questions can be solved by using just the basic R graphics functions.*

1. What are the 3 most fuel-efficient models in terms of highway mileage? Storage the data frame (with 3 rows) containing all information of these three cars in a variable named `top3` and print.
```{r}
library(ggplot2)
a<-c(0)
b<-c(0)
c<-c(0)
d<-c(0)
e<-c(0)
f<-c(0)
for(i in 1:234){
if(a<mpg$hwy[i]){
a<-c(mpg$hwy[i])
d<-c(i)
}
else if(b<mpg$hwy[i]){
b<-c(mpg$hwy[i])
e<-c(i)
}
else if(c<mpg$hwy[i]){
c<-mpg$hwy[i]
f<-c(i)
}
}
mpg[c(d,e,f),]

```
In this code vector a,b,c store the highest highway mileage and if throughout the dataframe it finds a highway mileage that is larger than a,b,or c it will overwrite it and update d,e,f which are the index value of a,b,c respectively. Once the program is over I then get the rows of mpg at the indexs labeled by d,e,f and display them. 

2. How many different compact models are included in this dataset? Storage the answer in  a variable named `numCompact` and print.
```{r}
numCompact=0
for(i in 1:234){
if(mpg$class[i]=='compact'){
numCompact<-numCompact+1
}
}
numCompact
```
I created a numbernumCompact at 0 and then throughout the entire dataframe if the class at mpg's row index is equal to the string compact will increment the number numCompact. At the end I display numCOmpact to see how many cars are compact.

3. Create a scatterplot of `hwy` and `cty` and describe the relationship. Why are there so few points visible? Explain your findings.
```{r}
plot(mpg$hwy,mpg$cty,main="hwy vs cty",xlab="hwy",ylab="cty")
```
I used the plot function for hwy and cty and found that there is a clear positive relationship. 

4. Are cars in 2008 more fuel efficient than cars in 1999? Use graphs *and* summary statistics to answer this question. Explain your findings.
```{r}
mpg$hwy
fuelEfficient<-c(0)
for(i in 1:234){
if(i==1){
fuelEfficient<-c((mpg$hwy[i]+mpg$cty[i])/2)
}
else{
fuelEfficient<-c(fuelEfficient,(mpg$hwy[i]+mpg$cty[i])/2)
}
}
plot(fuelEfficient,mpg$year)
mpg<-cbind(mpg,fuelEfficient)
df1999<-subset(mpg,mpg$year=="1999")
df2008<-subset(mpg,mpg$year=="2008")

summary(df2008)
summary(df1999)

```

Note: your submission is supposed to be fully reproducible, i.e. the TA and I will 'knit' your submission. Make sure that you don't use any file structure that depends on your computing environment.